Single-cell virus sequencing of influenza infections that trigger innate immunity

### 2019-04-26 paper/paper.tex
added DataDryad links for supp files, remove citations in supp

\caption{
Validation of reporter cell lines (\FIG{IFNrare}A) to identify IFN+ cells.
(A)
To validate the IFN reporter cell lines, they were infected at high MOI with the Cantell strain of Sendai virus, which strongly activates IFN expression~\cite{strahle2006sendai}.
The name of each of reporter cell line is indicated at the top of each row of plots.
At 13 hours post-infection, activation of the IFN reporter was then monitored by flow cytometry using the marker indicated at the bottom of each plot (either a fluorescent protein or antibody staining for the cell-surface LNGFR$\Delta$C using a PE-conjugated anti-LNGFR antibody from Miltenyi Biotec).
Sendai infection efficiently activated the IFN reporter in all cases, with the strongest signal from the IFN-$\lambda$ reporter driving ZsGreen.
(B)
The type I and type III IFN reporters are highly correlated in their activation.
An A549 cell line was generated by transduction with both the IFN-$\beta$ and IFN-$\lambda$ reporters driving expression of mCherry and ZsGreen, respectively.
The cells were then infected with two different stocks of ``wild-type'' WSN influenza, or stocks with a deletion in PB1 or stop codons in NS1 (described later in the paper).
After 13 hours, cells were analyzed by flow cytometry.
Cells positive for either fluorescent reporter were further analyzed.
As shown in the FACS plots, expression of the \textit{IFNB1} and \textit{IFNL1} reporters is highly correlated in all cases.
}
\label{suppfig:reportervalidation}
\end{suppfig}

\begin{suppfig}
\includegraphics[width=0.9\textwidth]{figures/IFN_stochastic/Flow/flow_plot.pdf}
\caption{
Flow cytometry data for \FIG{IFNrare}B.
The A549 cells with the \textit{IFNL1} reporter driving LNGFR$\Delta$C-ZsGreen were not infected, infected with saturating amounts of the Cantell strain of Sendai virus~\cite{strahle2006sendai}, or infected the same stock of influenza virus used in the single-cell experiment at a target MOI of 0.3.
After 13 hours, the cells were stained for expression of HA protein and analyzed by FACS for HA and expression of the ZsGreen driven by the \textit{IFNL1} reporter.
Each condition was done in triplicate.
The contour plots show the density of all cells, and all IFN+ cells are also indicated by orange dots.
Cells were classified as HA+ or IFN+ based on gates set to put 0.05\% of the uninfected cells in these populations.
For the influenza-infected cells, the percentage IFN+ was calculated only among the HA+ cells (since these are the ones that are infected).
For the uninfected and Sendai-virus infected, the percentage IFN+ was calculated among all cells, since these cells do not express HA.
}
\label{suppfig:IFNrare_flow}
\end{suppfig}

@@

\begin{suppfig}
\includegraphics[width=0.5\textwidth]{figures/single_cell_figures/p_frac_has_gene.pdf}
\caption{
The fraction of infected cells that are called as expressing each viral gene.
The gray dashed line is at one (the fraction that would be observed if all viral genes are expressed in all infected cells).
Each viral gene is detected in $\sim$80-90\% of the infected cells, roughly in line with prior estimates~\cite{brooke2013most, heldt2015single, dou2017analysis, russell2018extreme}.
The exception is NP, which is detected in virtually all infected cells.
The much higher frequency of detecting NP could reflect a biological phenomenon, but we suspect it is more likely that cells lacking NP tend to have much lower viral gene expression overall and so are not reliably called as being infected in our experiments because the number of viral mRNAs is below the detection limit.}
\label{suppfig:frac_has_gene}
\end{suppfig}

@@

\begin{suppfig}
\includegraphics[width=0.8\textwidth]{figures/single_cell_figures/p_isg.pdf}
\caption{
Expression of ISGs in single infected and uninfected cells.
For each cell, we quantified ISG expression as the total fraction of cellular mRNAs derived from four prototypical ISGs (IFIT1, ISG15, CCL5, and Mx1). 
(A) The histograms show the distribution of ISG expression taken across infected (top) and uninfected (bottom) cells.
We heuristically classify as ISG+ cells with $>10^{-3}$ of their cellular mRNA from ISGs, and color these cells red.
Comparison to \FIG{transcriptomics}G shows that substantially more cells are ISG+ than IFN+, both among infected and uninfected cells.
This is probably because paracrine signaling can induce ISG expression in cells that are not themselves expressing IFN~\cite{stetson2006type,honda2006type}.
(B) Correlation between the fraction of cellular mRNA derived from IFN and ISGs.
Each point represents one cell, and the Pearson correlation coefficient is shown.
IFN and ISG expression are more correlated for infected than uninfected cells, probably because in the latter the ISG expression is more often due to paracrine signaling that does not induce expression of IFN itself.
Among both the infected and uninfected populations, there are many cells with high expression of ISGs and little expression of IFN, but no cells that express high levels of IFN without also substantially expressing ISGs.
}
\label{suppfig:ISG}
\end{suppfig}

@@

\begin{suppfig}
\includegraphics[width=\textwidth]{figures/single_cell_figures/p_tsne.pdf}
\caption{
Unsupervised t-SNE clustering shows that cell-to-cell variation in expression of influenza, IFN, and ISG transcripts substantially contributes to the structure of the data.
To generate an unbiased representation of the factors that distinguished the transcriptomes of the cells in our experiments, we used unsupervised t-SNE clustering~\cite{maaten2008visualizing} as implemented in \texttt{Monocle}~\cite{qiu2017reversed, trapnell2014dynamics} to generate a two-dimensional representation of the data.
In the t-SNE plot, each point is a different cell, and cells with similar transcriptomes are closer together.
Each panel shows the same t-SNE plot, but the cells are colored differently in each panel based on the amount of viral, IFN, or ISG mRNA, shown on a log (top) or linear (bottom) scale.
As is clear from this plot, expression of influenza, IFN, and ISG genes contributes substantially to the structure of the data, since cells with high expression of these genes clearly group together.
}
\label{suppfig:tSNE}
\end{suppfig}

\begin{suppfig}
\includegraphics[width=\textwidth]{figures/pacbio_single_cell_figures/ccs_per_viralbarcode.pdf}
\caption{
The number of PacBio CCSs that passed quality-control steps and aligned to an influenza virus gene.
These sequences were obtained using several PacBio runs, most of which were intentionally loaded with different amounts of the various viral genes to increase coverage on genes that were needed in order to obtain the full sequences of virions infecting cells (see \SUPPFILE{pacbio_analysis}).
Because of this unequal loading and the inherently different PCR amplification efficiencies of different viral genes, unlike the transcriptomic data in \FIG{transcriptomics}, the numbers of CCSs for different genes should \emph{not} be taken as an indicator of their abundance in the infected cells.
Especially for the polymerase genes (PB2, PB1, and PA), many CCSs corresponded to genes with internal deletions, since these shorter forms of the genes were preferentially amplified during PCR.
Therefore, the plot is faceted by the number of CCSs for any length of the gene, and for full-length genes.
Note that the disproportionate sequencing of the shorter internally deleted genes does not greatly affect the genotype calling in \FIG{genotypes} since UMIs were used to collapse sequences derived from the same cDNA, and cell barcodes were used to collapse sequences from the same cell.
The bars in the plot are colored by whether the sequence is derived from the wild-type viral variant, the synonymously barcoded viral variant, or represents a mixed-barcode molecule (see panel B).
From the frequencies of these different forms, we estimate~\cite{bloom2018estimating} that 5.7\% of molecules are chimeric due to PCR strand exchange.
About half of these PCR chimeras could be identified by the presence of mixed viral barcodes and removed from subsequent analyses, leaving $\sim$3\% un-identified chimeras.
For some molecules (mostly polymerase genes with internal deletions) one of the barcode sites was deleted from the molecule and so the barcode identity could only be partially called.
A negligible number of molecules have low-accuracy sequence or unexpected nucleotide identities at the sites of the viral barcodes.
}
\label{suppfig:CCSs}
\end{suppfig}

\begin{suppfig}
\includegraphics[width=0.8\textwidth]{figures/StrandExchangeSchematic/StrandExchangeSchematic.pdf}
\caption{
Strategy for detecting strand exchange during sequencing of full-length viral genes.
The library preparation for PacBio sequencing of the cDNA for the full-length viral genes required many cycles of PCR.
A major concern is that strand exchange during this PCR could scramble mutations and 10X cell barcodes / UMIs from different molecules.
We detect PCR strand exchange by leveraging the fact that our cells were infected with a mix of wild-type virus and virus carrying synonymous barcodes near both termini of each gene.
If there is no strand exchange, all molecules should either be wild-type or have the synonymous barcoding mutations at \emph{both} termini.
Strand exchange will create some molecules that have wild-type nucleotides at one termini and synonymous barcoding mutations at the other termini.
\SUPPFIG{CCSs} shows the frequencies with which these different types of molecules were observed during the PacBio sequencing.
Note that since the rate of homologous recombination in influenza virus in negligible~\cite{boni2008homologous}, such mixed-barcode molecules are \emph{not} expected to be generated naturally during co-infection.
}
\label{suppfig:StrandExchange}
\end{suppfig}

@@

\begin{suppfig}
{\bf \Large A}\includegraphics[width=0.48\textwidth, valign=t]{figures/Validation_Figure/SNP_flow_plot.pdf}
\hspace{0.02\textwidth}
{\bf \Large B} \includegraphics[width=0.39\textwidth, valign=t]{figures/Validation_Figure/del_flow_plot.pdf}
\caption{
Flow cytometry data for \FIG{validation}.
(A)
Data for \FIG{validation}A.
A549 cells with the \textit{IFNL1} reporter driving LNGFR$\Delta$C-ZsGreen were infected with stocks of the indicated mutant.
After 13 hours, cells were stained for HA protein and analyzed by FACS.
Contour plots show density of all cells, and IFN+ cells are also indicated by orange dots.
Cells were classified as HA+ or IFN+ based on gates set to put 0.05\% of uninfected cells in these populations.
For infected cells, the percentage IFN+ was calculated among the HA+ cells (since these are the ones that are infected).
For uninfected cells, the percentage IFN+ was calculated among all cells, since uninfected cells do not express HA.
For each viral mutant, two independent stocks were assayed in duplicate (i.e., \#1a and \#1b are one viral stock, and \#2a and \#2b are the other).
The infections with replicate \#1 of the wild-type virus were performed at an MOI of 0.1 as determined by TCID50, and all other viruses were infected at an equivalent particle number as determined by HI assay. 
(B)
Data for \FIG{validation}B.
The virus with the deletion in PB1 cannot be normalized by HA expression since it expresses less HA due to the lack of secondary transcription.
Therefore, all cells were infected at an equivalent MOI of 0.3 as determined by TCID50 on MDCK-SIAT1 cells for wild type and NS1stop, and on MDCK-SIAT1 cells expressing PB1~\cite{bloom2010permissive} for PB1del385to2163.
\SUPPFIG{delqPCR} shows that at these equivalent TCID50s, all variants had similar amounts of transcriptionally active virus in the absence of secondary transcription.
The percent IFN+ was calculated for \emph{all} cells (HA+ and HA-) since that is a more fair comparison for PB1del385to2163.
}
\label{suppfig:validationflow}
\end{suppfig}

\begin{suppfig}
\includegraphics[width=0.6\textwidth]{figures/Validation_Figure/qpcr_plot.pdf}
\caption{
Validation that the infections in \FIG{validation}B and \SUPPFIG{validationflow}B were performed at similar doses of virions capable of initiating primary transcription.
In this experiment, A549 cells were infected at MOI of 0.4 (based on TCID50 as described in \SUPPFIG{validationflow}B), and then after 8 hours mRNA was harvested for qPCR on oligo-dT primed reverse transcription products.
The y-axis shows the ratio of viral HA mRNA to the housekeeping gene L32.
These infections were performed in the presence of absence of 50 $\mu$g/ml cycloheximide, which blocks protein synthesis and hence secondary transcription by newly synthesized viral proteins~\cite{killip2014activation}.
In the absence of cycloheximide, the viruses with deletions in PB1 produced less viral mRNA presumably because they could not produce PB1 protein for secondary transcription.
But in the presence of cycloheximide, all viruses produced similar amounts of viral mRNA, indicating that the dose of particles active for primary transcription is roughly equivalent across variants.
Each measurement was performed in quadruplicate.
}
\label{suppfig:delqPCR}
\end{suppfig}

@@

\begin{suppfile}
\caption{
Sequences of the IFN reporters in \FIG{IFNrare}A are at \url{https://github.com/jbloomlab/IFNsorted_flu_single_cell/tree/master/paper/figures/IFN_stochastic/IFN_reporter/plasmids}.
}
\label{suppfile:reporter_sequences}
\end{suppfile}

\begin{suppfile}
\caption{
Genbank files giving sequences of the wild-type and synonymously barcoded viruses are at \url{https://github.com/jbloomlab/IFNsorted_flu_single_cell/blob/master/data/flu_sequences/flu-wsn.gb} and \url{https://github.com/jbloomlab/IFNsorted_flu_single_cell/blob/master/data/flu_sequences/flu-wsn-double-syn.gb}.
}
\label{suppfile:virus_seqs}
\end{suppfile}

\begin{suppfile}
\caption{
A text file giving the primers used to amplify the influenza cDNAs for PacBio sequencing is at \url{https://github.com/jbloomlab/IFNsorted_flu_single_cell/tree/master/paper/figures/WorkflowSchematic/PacBio_primer_list.txt} 
}
\label{suppfile:primer_sequences}
\end{suppfile}

\begin{suppfile}
\caption{
A CSV file giving the genotypes in \FIG{genotypes} is at \url{https://github.com/jbloomlab/IFNsorted_flu_single_cell/blob/master/paper/figures/single_cell_figures/genotypes.csv}.
}
\label{suppfile:genotypes}
\end{suppfile}

\begin{suppfile}
\caption{
A CSV file giving the viral mutations and related information in \FIG{mutations} is at \url{https://github.com/jbloomlab/IFNsorted_flu_single_cell/blob/master/paper/figures/single_cell_figures/mutations.csv}.
}
\label{suppfile:mutations}
\end{suppfile}

\begin{suppfile}
\caption{
Genbank plasmid maps for the mutant genes cloned into the pHW* bi-directional reverse genetics plasmid~\cite{hoffmann2000dna} are at \url{https://github.com/jbloomlab/IFNsorted_flu_single_cell/tree/master/paper/figures/FluVariantPlasmidMaps}.
}
\label{suppfile:validation_sequences}
\end{suppfile}

\begin{suppfile}
\caption{\label{suppfile:pacbio_analysis}
The Jupyter notebook that analyzes the PacBio data is at \url{https://github.com/jbloomlab/IFNsorted_flu_single_cell/blob/master/pacbio_analysis.ipynb}.
}
\end{suppfile}

\begin{suppfile}
\caption{\label{suppfile:monocle_analysis}
The Jupyter notebook that analyzes the annotated cell-gene matrix is at \url{https://github.com/jbloomlab/IFNsorted_flu_single_cell/blob/master/monocle_analysis.ipynb}.
}
\end{suppfile}

>>>>

\caption{
Validation of reporter cell lines (\FIG{IFNrare}A) to identify IFN+ cells.
(A)
To validate the IFN reporter cell lines, they were infected at high MOI with the Cantell strain of Sendai virus, which strongly activates IFN expression.
The name of each of reporter cell line is indicated at the top of each row of plots.
At 13 hours post-infection, activation of the IFN reporter was then monitored by flow cytometry using the marker indicated at the bottom of each plot (either a fluorescent protein or antibody staining for the cell-surface LNGFR$\Delta$C using a PE-conjugated anti-LNGFR antibody from Miltenyi Biotec).
Sendai infection efficiently activated the IFN reporter in all cases, with the strongest signal from the IFN-$\lambda$ reporter driving ZsGreen.
(B)
The type I and type III IFN reporters are highly correlated in their activation.
An A549 cell line was generated by transduction with both the IFN-$\beta$ and IFN-$\lambda$ reporters driving expression of mCherry and ZsGreen, respectively.
The cells were then infected with two different stocks of ``wild-type'' WSN influenza, or stocks with a deletion in PB1 or stop codons in NS1 (described later in the paper).
After 13 hours, cells were analyzed by flow cytometry.
Cells positive for either fluorescent reporter were further analyzed.
As shown in the FACS plots, expression of the \textit{IFNB1} and \textit{IFNL1} reporters is highly correlated in all cases.
}
\label{suppfig:reportervalidation}
\end{suppfig}

\begin{suppfig}
\includegraphics[width=0.9\textwidth]{figures/IFN_stochastic/Flow/flow_plot.pdf}
\caption{
Flow cytometry data for \FIG{IFNrare}B.
The A549 cells with the \textit{IFNL1} reporter driving LNGFR$\Delta$C-ZsGreen were not infected, infected with saturating amounts of the Cantell strain of Sendai virus, or infected the same stock of influenza virus used in the single-cell experiment at a target MOI of 0.3.
After 13 hours, the cells were stained for expression of HA protein and analyzed by FACS for HA and expression of the ZsGreen driven by the \textit{IFNL1} reporter.
Each condition was done in triplicate.
The contour plots show the density of all cells, and all IFN+ cells are also indicated by orange dots.
Cells were classified as HA+ or IFN+ based on gates set to put 0.05\% of the uninfected cells in these populations.
For the influenza-infected cells, the percentage IFN+ was calculated only among the HA+ cells (since these are the ones that are infected).
For the uninfected and Sendai-virus infected, the percentage IFN+ was calculated among all cells, since these cells do not express HA.
}
\label{suppfig:IFNrare_flow}
\end{suppfig}

@@

\begin{suppfig}
\includegraphics[width=0.5\textwidth]{figures/single_cell_figures/p_frac_has_gene.pdf}
\caption{
The fraction of infected cells that are called as expressing each viral gene.
The gray dashed line is at one (the fraction that would be observed if all viral genes are expressed in all infected cells).
Each viral gene is detected in $\sim$80-90\% of the infected cells.
The exception is NP, which is detected in virtually all infected cells.
The much higher frequency of detecting NP could reflect a biological phenomenon, but we suspect it is more likely that cells lacking NP tend to have much lower viral gene expression overall and so are not reliably called as being infected in our experiments because the number of viral mRNAs is below the detection limit.}
\label{suppfig:frac_has_gene}
\end{suppfig}

@@

\begin{suppfig}
\includegraphics[width=0.8\textwidth]{figures/single_cell_figures/p_isg.pdf}
\caption{
Expression of ISGs in single infected and uninfected cells.
For each cell, we quantified ISG expression as the total fraction of cellular mRNAs derived from four prototypical ISGs (IFIT1, ISG15, CCL5, and Mx1). 
(A) The histograms show the distribution of ISG expression taken across infected (top) and uninfected (bottom) cells.
We heuristically classify as ISG+ cells with $>10^{-3}$ of their cellular mRNA from ISGs, and color these cells red.
Comparison to \FIG{transcriptomics}G shows that substantially more cells are ISG+ than IFN+, both among infected and uninfected cells.
This is probably because paracrine signaling can induce ISG expression in cells that are not themselves expressing IFN.
(B) Correlation between the fraction of cellular mRNA derived from IFN and ISGs.
Each point represents one cell, and the Pearson correlation coefficient is shown.
IFN and ISG expression are more correlated for infected than uninfected cells, probably because in the latter the ISG expression is more often due to paracrine signaling that does not induce expression of IFN itself.
Among both the infected and uninfected populations, there are many cells with high expression of ISGs and little expression of IFN, but no cells that express high levels of IFN without also substantially expressing ISGs.
}
\label{suppfig:ISG}
\end{suppfig}

@@

\begin{suppfig}
\includegraphics[width=\textwidth]{figures/single_cell_figures/p_tsne.pdf}
\caption{
Unsupervised t-SNE clustering shows that cell-to-cell variation in expression of influenza, IFN, and ISG transcripts substantially contributes to the structure of the data.
To generate an unbiased representation of the factors that distinguished the transcriptomes of the cells in our experiments, we used unsupervised t-SNE clustering as implemented in \texttt{Monocle} to generate a two-dimensional representation of the data.
In the t-SNE plot, each point is a different cell, and cells with similar transcriptomes are closer together.
Each panel shows the same t-SNE plot, but the cells are colored differently in each panel based on the amount of viral, IFN, or ISG mRNA, shown on a log (top) or linear (bottom) scale.
As is clear from this plot, expression of influenza, IFN, and ISG genes contributes substantially to the structure of the data, since cells with high expression of these genes clearly group together.
}
\label{suppfig:tSNE}
\end{suppfig}

\begin{suppfig}
\includegraphics[width=\textwidth]{figures/pacbio_single_cell_figures/ccs_per_viralbarcode.pdf}
\caption{
The number of PacBio CCSs that passed quality-control steps and aligned to an influenza virus gene.
These sequences were obtained using several PacBio runs, most of which were intentionally loaded with different amounts of the various viral genes to increase coverage on genes that were needed in order to obtain the full sequences of virions infecting cells (see \SUPPFILE{pacbio_analysis}).
Because of this unequal loading and the inherently different PCR amplification efficiencies of different viral genes, unlike the transcriptomic data in \FIG{transcriptomics}, the numbers of CCSs for different genes should \emph{not} be taken as an indicator of their abundance in the infected cells.
Especially for the polymerase genes (PB2, PB1, and PA), many CCSs corresponded to genes with internal deletions, since these shorter forms of the genes were preferentially amplified during PCR.
Therefore, the plot is faceted by the number of CCSs for any length of the gene, and for full-length genes.
Note that the disproportionate sequencing of the shorter internally deleted genes does not greatly affect the genotype calling in \FIG{genotypes} since UMIs were used to collapse sequences derived from the same cDNA, and cell barcodes were used to collapse sequences from the same cell.
The bars in the plot are colored by whether the sequence is derived from the wild-type viral variant, the synonymously barcoded viral variant, or represents a mixed-barcode molecule (see panel B).
From the frequencies of these different forms, we estimate that 5.7\% of molecules are chimeric due to PCR strand exchange.
About half of these PCR chimeras could be identified by the presence of mixed viral barcodes and removed from subsequent analyses, leaving $\sim$3\% un-identified chimeras.
For some molecules (mostly polymerase genes with internal deletions) one of the barcode sites was deleted from the molecule and so the barcode identity could only be partially called.
A negligible number of molecules have low-accuracy sequence or unexpected nucleotide identities at the sites of the viral barcodes.
}
\label{suppfig:CCSs}
\end{suppfig}

\begin{suppfig}
\includegraphics[width=0.8\textwidth]{figures/StrandExchangeSchematic/StrandExchangeSchematic.pdf}
\caption{
Strategy for detecting strand exchange during sequencing of full-length viral genes.
The library preparation for PacBio sequencing of the cDNA for the full-length viral genes required many cycles of PCR.
A major concern is that strand exchange during this PCR could scramble mutations and 10X cell barcodes / UMIs from different molecules.
We detect PCR strand exchange by leveraging the fact that our cells were infected with a mix of wild-type virus and virus carrying synonymous barcodes near both termini of each gene.
If there is no strand exchange, all molecules should either be wild-type or have the synonymous barcoding mutations at \emph{both} termini.
Strand exchange will create some molecules that have wild-type nucleotides at one termini and synonymous barcoding mutations at the other termini.
\SUPPFIG{CCSs} shows the frequencies with which these different types of molecules were observed during the PacBio sequencing.
Note that since the rate of homologous recombination in influenza virus in negligible, such mixed-barcode molecules are \emph{not} expected to be generated naturally during co-infection.
}
\label{suppfig:StrandExchange}
\end{suppfig}

@@

\begin{suppfig}
{\bf \Large A}\includegraphics[width=0.48\textwidth, valign=t]{figures/Validation_Figure/SNP_flow_plot.pdf}
\hspace{0.02\textwidth}
{\bf \Large B} \includegraphics[width=0.39\textwidth, valign=t]{figures/Validation_Figure/del_flow_plot.pdf}
\caption{
Flow cytometry data for \FIG{validation}.
(A)
Data for \FIG{validation}A.
A549 cells with the \textit{IFNL1} reporter driving LNGFR$\Delta$C-ZsGreen were infected with stocks of the indicated mutant.
After 13 hours, cells were stained for HA protein and analyzed by FACS.
Contour plots show density of all cells, and IFN+ cells are also indicated by orange dots.
Cells were classified as HA+ or IFN+ based on gates set to put 0.05\% of uninfected cells in these populations.
For infected cells, the percentage IFN+ was calculated among the HA+ cells (since these are the ones that are infected).
For uninfected cells, the percentage IFN+ was calculated among all cells, since uninfected cells do not express HA.
For each viral mutant, two independent stocks were assayed in duplicate (i.e., \#1a and \#1b are one viral stock, and \#2a and \#2b are the other).
The infections with replicate \#1 of the wild-type virus were performed at an MOI of 0.1 as determined by TCID50, and all other viruses were infected at an equivalent particle number as determined by HI assay. 
(B)
Data for \FIG{validation}B.
The virus with the deletion in PB1 cannot be normalized by HA expression since it expresses less HA due to the lack of secondary transcription.
Therefore, all cells were infected at an equivalent MOI of 0.3 as determined by TCID50 on MDCK-SIAT1 cells for wild type and NS1stop, and on MDCK-SIAT1 cells expressing PB1 for PB1del385to2163.
\SUPPFIG{delqPCR} shows that at these equivalent TCID50s, all variants had similar amounts of transcriptionally active virus in the absence of secondary transcription.
The percent IFN+ was calculated for \emph{all} cells (HA+ and HA-) since that is a more fair comparison for PB1del385to2163.
}
\label{suppfig:validationflow}
\end{suppfig}

\begin{suppfig}
\includegraphics[width=0.6\textwidth]{figures/Validation_Figure/qpcr_plot.pdf}
\caption{
Validation that the infections in \FIG{validation}B and \SUPPFIG{validationflow}B were performed at similar doses of virions capable of initiating primary transcription.
In this experiment, A549 cells were infected at MOI of 0.4 (based on TCID50 as described in \SUPPFIG{validationflow}B), and then after 8 hours mRNA was harvested for qPCR on oligo-dT primed reverse transcription products.
The y-axis shows the ratio of viral HA mRNA to the housekeeping gene L32.
These infections were performed in the presence of absence of 50 $\mu$g/ml cycloheximide, which blocks protein synthesis and hence secondary transcription by newly synthesized viral proteins.
In the absence of cycloheximide, the viruses with deletions in PB1 produced less viral mRNA presumably because they could not produce PB1 protein for secondary transcription.
But in the presence of cycloheximide, all viruses produced similar amounts of viral mRNA, indicating that the dose of particles active for primary transcription is roughly equivalent across variants.
Each measurement was performed in quadruplicate.
}
\label{suppfig:delqPCR}
\end{suppfig}

@@

\begin{suppfile}
\caption{
Sequences of the IFN reporters in \FIG{IFNrare}A are at \url{https://github.com/jbloomlab/IFNsorted_flu_single_cell/tree/master/paper/figures/IFN_stochastic/IFN_reporter/plasmids}.
This file is also available on Data Dryad Digital Repository at \url{https://doi.org/10.5061/dryad.nh053c6}.
}
\label{suppfile:reporter_sequences}
\end{suppfile}

\begin{suppfile}
\caption{
Genbank files giving sequences of the wild-type and synonymously barcoded viruses are at \url{https://github.com/jbloomlab/IFNsorted_flu_single_cell/blob/master/data/flu_sequences/flu-wsn.gb} and \url{https://github.com/jbloomlab/IFNsorted_flu_single_cell/blob/master/data/flu_sequences/flu-wsn-double-syn.gb}.
This file is also available on Data Dryad Digital Repository at \url{https://doi.org/10.5061/dryad.nh053c6}.
}
\label{suppfile:virus_seqs}
\end{suppfile}

\begin{suppfile}
\caption{
A text file giving the primers used to amplify the influenza cDNAs for PacBio sequencing is at \url{https://github.com/jbloomlab/IFNsorted_flu_single_cell/tree/master/paper/figures/WorkflowSchematic/PacBio_primer_list.txt}.
This file is also available on Data Dryad Digital Repository at \url{https://doi.org/10.5061/dryad.nh053c6}. 
}
\label{suppfile:primer_sequences}
\end{suppfile}

\begin{suppfile}
\caption{
A CSV file giving the genotypes in \FIG{genotypes} is at \url{https://github.com/jbloomlab/IFNsorted_flu_single_cell/blob/master/paper/figures/single_cell_figures/genotypes.csv}.
This file is also available on Data Dryad Digital Repository at \url{https://doi.org/10.5061/dryad.nh053c6}.
}
\label{suppfile:genotypes}
\end{suppfile}

\begin{suppfile}
\caption{
A CSV file giving the viral mutations and related information in \FIG{mutations} is at \url{https://github.com/jbloomlab/IFNsorted_flu_single_cell/blob/master/paper/figures/single_cell_figures/mutations.csv}.
This file is also available on Data Dryad Digital Repository at \url{https://doi.org/10.5061/dryad.nh053c6}.
}
\label{suppfile:mutations}
\end{suppfile}

\begin{suppfile}
\caption{
Genbank plasmid maps for the mutant genes cloned into the pHW* bi-directional reverse genetics plasmid are at \url{https://github.com/jbloomlab/IFNsorted_flu_single_cell/tree/master/paper/figures/FluVariantPlasmidMaps}.
This file is also available on Data Dryad Digital Repository at \url{https://doi.org/10.5061/dryad.nh053c6}.
}
\label{suppfile:validation_sequences}
\end{suppfile}

\begin{suppfile}
\caption{\label{suppfile:pacbio_analysis}
The Jupyter notebook that analyzes the PacBio data is at \url{https://github.com/jbloomlab/IFNsorted_flu_single_cell/blob/master/pacbio_analysis.ipynb}.
This file is also available on Data Dryad Digital Repository at \url{https://doi.org/10.5061/dryad.nh053c6}.
}
\end{suppfile}

\begin{suppfile}
\caption{\label{suppfile:monocle_analysis}
The Jupyter notebook that analyzes the annotated cell-gene matrix is at \url{https://github.com/jbloomlab/IFNsorted_flu_single_cell/blob/master/monocle_analysis.ipynb}.
This file is also available on Data Dryad Digital Repository at \url{https://doi.org/10.5061/dryad.nh053c6}.
}
\end{suppfile}
